---
title: 'MSDS 6306 Assignment 09: Web Scrapping'
author: "J Lancon"
date: "November 6, 2017"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    fig_caption: yes
---
### Assignment: Unit 09 Assignment
#### Course: MSDS 6306 - 405 Doing Data Science
#### Project Description: Web Scrapping
##### GitHub Repository: https://github.com/jlancon/MSDS6306/tree/master/Homework/Unit_9

&nbsp;&nbsp;&nbsp; **Problem #1:** Navigate to IMDB website and find the movie [Harry Potter and the Deathly Hallows: Part 2](http://www.imdb.com/title/tt1201607/fullcredits?ref_=tt_ql_1).  Scrape the page with any R package you are comfortable with. We are interested is the table of the Cast in order of crediting.

* Parse down the table to only include the Actor's Name and their corresponding Character in the movie.  
* Clean observation; eliminating empty fields and fields with improper data.  
* One Actor plays two characters, whos character name(s) has special(\n) characters in its description.  Replace his character name, removing special characters, so it displays properly.  
* The Actor variable includes first name and surname.  Split the variable into FirstName and Surname.  Play special attention to multiple names/middle names, etc.  
* Display the first 10 rows of the 'cleaned' dataframe with 3 variables (FirstName, Surname, Character)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
library(knitr)
library(stringr)
library(rvest)
library(ggplot2)
```
##### Reading Website / Scrapping Data
&nbsp;&nbsp;&nbsp;**Reading website** into R
Website Address for movie: http://www.imdb.com/title/tt1201607/fullcredits?ref_=tt_ql_1  

```{r ReadingHtml}
URLMovie <- read_html("http://www.imdb.com/title/tt1201607/fullcredits?ref_=tt_ql_1")
```
&nbsp;&nbsp;&nbsp;**Scrapping Table** with Actor & Character information
```{r ScrappingTable}
# From YouTube video https://www.youtube.com/watch?v=gSbuwYdNYLM
Table_Data <- URLMovie %>%
      html_nodes('table') %>% .[[3]] %>%
      html_table()
```
&nbsp;&nbsp;&nbsp;**Parsing Table** to only include Actor's Full Name and corresponding Character in movie.
```{r ParsingTable}
Actor_Character.df <- Table_Data[,c(2,4)]
colnames(Actor_Character.df) <- c("ActorFullName","Character")
```
&nbsp;&nbsp;&nbsp;**Cleaning erroneous observations** and modifying Character observation for Actor who plays multiple roles(eliminating special character(s))  
```{r CleaningData}
Actor_Character.df <- Actor_Character.df[- grep("Rest of cast listed",Actor_Character.df$ActorFullName),]
Actor_Character.df <- Actor_Character.df[!apply(Actor_Character.df == "",1,all),]
Character.df <- as.data.frame(sapply(Actor_Character.df[,2],gsub,pattern=".*Grip.*",replacement="Griphook / Professor Filius Flitwick"))
```
&nbsp;&nbsp;&nbsp;**Splitting Actor Name Field** into First Name and Surname, using technique that will leave Middle names and/or multiple first names in First Name field.  

```{r SplittingName}
First_Surname.df <- as.data.frame(str_split_fixed(Actor_Character.df$ActorFullName, ' (?=[^ ]+$)',2))
#Alternate Solution
#First_Surname.df <- as.data.frame(do.call(rbind, strsplit(Actor_Character_Data$Actor, ' (?=[^ ]+$)', perl=TRUE)))
```
&nbsp;&nbsp;&nbsp;**Combining Dataframes** into Cleaned Dataset for use in future analysis.  
&nbsp;&nbsp;&nbsp;**Displaying Table** for reference. Only displaying first 10 observations.

```{r TableDisplay}
FinalList.df <- cbind(First_Surname.df,Character.df)
colnames(FinalList.df) <- c("FirstName", "Surname", "Character")

kable(head(FinalList.df,10), align="c", caption="Harry Potter and the Deathly Hallows: Part 2 Cast and Character",format="markdown",row.names = FALSE)
```
: **Harry Potter and the Deathly Hallows Part 2:  Cast and Characters**

***


#########################################################
#########################################################
&nbsp;&nbsp;&nbsp; **Problem #2:** Navigate to ESPN website, there are statistics of each NBA player, and find the [San Antonio Spurs stats webpage](http://www.espn.com/nba/team/stats/_/name/sa/san-antonio-spurs); which displays player's current statistics.

* Interested in **Shooting Statistics** table.
* Parse down the table to only include the Player's Name and Field Goal Percentage (FG%).  
* Clean observation; eliminating empty fields and fields with improper data.  
* One Actor plays two characters, whos character name(s) has special(\n) characters in its description.  Replace his character name, removing special characters, so it displays properly.  
* The Player catagory includes the Full Name of player as well as position.  Split the variable into Player Name and Position.  
* Create a 'cleaned' dataset with Player's Name, Position, and Field Goal Percentage
* Create a bar graph displaying Player and their FG%.  Color the columns by the Player's position.  

##### Reading Website / Scrapping Data
&nbsp;&nbsp;&nbsp;**Reading website** into R
Website Address for Team stats: http://www.espn.com/nba/team/stats/_/name/sa/san-antonio-spurs  
```{r SpursURL}
URLSpurs <- read_html('http://www.espn.com/nba/team/stats/_/name/sa/san-antonio-spurs')
```

&nbsp;&nbsp;&nbsp;**Scrapping Table** with Player and Statistics information
```{r ScrapTableSpurs}
Table_Data_Spurs <- URLSpurs %>%
  html_nodes('table') %>% .[[2]] %>%
  html_table()
```

&nbsp;&nbsp;&nbsp;**Parsing Table** to only include Player's Name and FG% statistics.  
```{r SpurParsing}
Player_FGP.df <- Table_Data_Spurs[,c(1,4)]
```

&nbsp;&nbsp;&nbsp;**Cleaning Erroneous Observations** 
```{r CleaningSpurs}
Player_FGP.df <- Player_FGP.df[- grep("Totals",Player_FGP.df$X1),]
Player_FGP.df <- Player_FGP.df[- grep("SHOOTING STATISTICS",Player_FGP.df$X1),]
Player_FGP.df <- Player_FGP.df[- grep("PLAYER",Player_FGP.df$X1),]
```

&nbsp;&nbsp;&nbsp;**Extracting Variable and Changing Data Class** of FG% from Character to Numeric. 
```{r FGPExtraction}
FGPercent.df <- as.data.frame(as.numeric(Player_FGP.df$X4))
```

&nbsp;&nbsp;&nbsp;**Splitting Player's Name Variable** into Player's Name and Position. Eliminate unnecessary characters "," in variable observations.
```{r PlayerSplitting}
Player_Position.df <- as.data.frame(str_split_fixed(Player_FGP.df$X1, ' (?=[^ ]+$)',2))
Player_Position.df <- as.data.frame(sapply(Player_Position.df,gsub,pattern=",",replacement=""))
```

&nbsp;&nbsp;&nbsp;**Combining Dataframes** into Cleaned Dataset with Player's Name, Position, and FG%.  
```{r FinalSpurs}
Final.df <- cbind(Player_Position.df,FGPercent.df)
colnames(Final.df) <- c("Player", "Position", "FG_Percent")
Final.df <- Final.df[order(Final.df$FG_Percent, decreasing = TRUE), ]
```

&nbsp;&nbsp;&nbsp;**SA Spurs Bar Graph** displaying Player's current Field Goal Percentages for 2017/18 season.
```{r barGraphSpurs}
ggplot(data=Final.df, aes(x=reorder(Player, order(Final.df$FG_Percent, decreasing=FALSE)),y=(FG_Percent*100),fill=Position)) +
  geom_bar(position="dodge",stat="identity") + 
  scale_fill_brewer(palette="Spectral") +
#  scale_fill_hue(c=80, l=60) +
  coord_flip() +
  xlab("Players") +
  ylab("Field Goal Percentage") +
  ggtitle("San Antonio Spurs 2017/18\n Field Goal Percentages") +
  theme(plot.title = element_text(hjust = 0.5, face="bold", size=14)) +
  theme(axis.title = element_text(face="bold", size=10)) +
  theme(axis.title.y = element_text(face="bold", size=10)) +
  theme(panel.border = element_rect(linetype = "solid", fill = NA)) +
  theme(plot.margin = unit(c(1,1,1,1), "cm"))
```
